### 2023-10-03 main.tex
fixed

In this section we aim to use duality and the matrix product ansatz to determine the non equilibrium steady state and its arbitrary correlation for the integrable version of the multi-species stirring process, defined in Subsection \ref{integrableChain-duality}. As it will turn out, duality simplifies the commutation relations of the matrix product ansatz allowing to determined a closed formula for all correlations. As a by-product, absorption probabilities can be explicitly determined, giving a further probabilistic interpretation for the non-equilibrium steady state. 
\subsection{The integrable chain and the matrix product ansatz}

@@

 the Master equation
    \frac{d|{\Psi}(t)\rangle}{dt}=H|{\Psi}(t)\rangle

@@

	H|\Psi\rangle =0

@@

	Z_{L}=\langle V^{*}|(X_{1}+\ldots +X_{N})^{L}|V\rangle
where the operators $(X_{a})_{a\in \{1,\ldots,N\}}$ fulfill the commutators

@@

	\langle V^{*}|\left(\alpha_{a}(X_{1}+\ldots+X_{N})-X_{a}\right)=\lambda_{a}\langle V^{*}|\qquad\forall a=1,\ldots,N

@@

	\left(\beta_{a}(X_{1}+\ldots+X_{N})-X_{a}\right)|V\rangle=-\lambda_{a}|V\rangle\qquad\forall a=1,\ldots,N

@@

\textbf{Remark}: only $N-1$ of the $N$ equations \eqref{leftBoundary}  are independent. This can be seen by summing them over the species $a\in \{1,\ldots,N\}$:
		\sum_{a=1}^{N}	\langle V^{*}|\left(\alpha_{a}(X_{1}+\ldots+X_{N})-X_{a}\right)=\sum_{a=1}^{N}\lambda_{a}\langle V^{*}|=0

@@

The MPA gives an abstract form of the steady state in terms of an algebra of operators. However, to write explicitly the steady state one should provide a representation of this algebra. Another approach is to directly use the commutators and the boundary actions to compute correlations. This path has been followed in \cite{vanicat2017exact} to find one-point and two-point correlations. 
Here we follow the same approach. We will show that the duality property established in Section \ref{sectionDuality} simplifies substantially the  use of commutators in the

@@

where $H^{'}= S_{1} H S_{1}^{-1} $ has both boundary terms in a triangular form, and $H^{''} = S_{2}H' S_{2}^{-1}$ has the left boundary in a triangular form and the right boundary is diagonal. Thanks to these transformations the commutators \eqref{bulk} and the action on the boundary vectors \eqref{leftBoundary}, \eqref{rightBoundary} of the operators $(X_{a})_{a\in\{1,\ldots,N\}}$ of the MPA become simpler, therefore, the right eigenvector with vanishing eigenvalue $|\Psi^{''}\rangle$ of $H^{''}$ will be written explicitly. Finally, reversing the transformations we will retrieve $|\Psi\rangle$.

The first transformation $S_{1}$ is closely related to the dual matrix and we will see that $H^{'}$ is $\widetilde{H}^{T}$, up to the extra-site term described by the bosonic creation operator in \eqref{siteDualHamiltonian}. Therefore, $H^{'}$ is not stochastic. The second transformation $S_{2}$ diagonalize the right boundary. Finally, the composition  $S_{2} S_{1}$ allows to simplify the commutators of the MPA algebra. \\

The computation are organized as follows: we first define the similarity transformations $S_{1},S_{2}$ and the Hamiltonians $H^{'},H^{''}$. Second, we compute right eigenvector with vanishing eigenvalue of $H^{''}$ thanks to the simpler commutators. Finally, we invert the similarities and retrieve the right eigenvector with vanishing eigenvalue of $H^{'}$ and  the arbitrary correlations in the non-equilibrium steady state. The explicit expression of the steady state $|\Psi\rangle$ of $H$ is reported in Appendix \ref{appB}.

@@

	\mathcal{S}_{1}:=\exp{\left(\sum_{a=1}^{N-1}e_{Na}\right)}\label{S1Gen}\\
	\mathcal{S}_{2}:=\exp{\left(-\sum_{a=1}^{N-1}\beta_{a}e_{aN}\right)}\label{S2Gen}

@@

 \mathcal{S}_{1}^{x}:=\exp{\left(\sum_{a=1}^{N-1}e_{Na}^{x}\right)}%=\mathbbm{1}+\sum_{a=2}^{N}e_{1a}

@@

 $S_{1}$ is invertible, then we introduce 

@@

   H^{'}:=H^{'}_{left}+H_{bulk}+H^{'}_{L}

@@

	H^{'}_{left}=\sum_{a=1}^{N-1}\left(\alpha_{a}e_{aN}^{1}-e_{aa}^{1}\right)\qquad H^{'}_{L}=\sum_{a=1}^{N-1}\left(\beta_{a}e_{aN}^{L}-e_{aa}^{L}\right)

@@

		\sum_{a=1}^{N-1}\left(\alpha_{a}e_{aN}^{1}-e_{aa}^{1}\right)

@@

H_{left}^{'}=\exp{\left(\sum_{c=1}^{N-1}e_{1N}^{1}\right)}\sum_{a,b=1}^{N}\alpha_{a}\left(e_{ab}^{1}-e_{bb}^{1}\right)\exp{\left(-\sum_{c=1}^{N-1}e_{Nc}^{1}\right)}=\sum_{a=1}^{N-1}\left(\beta_{a}e_{aN}^{L}-e_{aa}^{L}\right)
\textbf{Remark:} the boundary Hamiltonians $H_{left}^{'}$ and $H_{right}^{'}$ relates to the transposed of  $\widetilde{H}_{left}$ and $\widetilde{H}_{right}$, i.e. to the transposed of the boundary part of the dual Hamiltonian  defined in \eqref{siteDualHamiltonian}. They are obtained replacing the  extra-site bosonic creation operators $(a^{\dagger})_{a}^{0}$  (resp. $(a^{\dagger})_{a}^{L+1}$) with the corresponding reservoir parameters $\alpha_a$ (resp. $\beta_a$).

@@

Obviously

@@

By taking the product over the chain

@@

	S_{2}=\prod_{x=1}^{L}S_{2}^{x}

@@

	H^{''}=S_{2}H^{'}S_{2}^{-1}

@@

	H_{left}^{''}=e_{NN}^{1}-\mathbbm{1}^{1}+\sum_{a=1}^{N-1}(\alpha_{a}-\beta_{a})e_{aN}^{1}\qquad H_{right}^{''}=e_{NN}^{L}-\mathbbm{1}^{L}
i.e. the left one is lower triangular and depends on the differences between boundary parameters, while the right one is diagonal. The expressions of $H_{left}^{''}$ and $H_{right}^{''}$ are proved using the Hadamard formula \eqref{HadamardFormula} with $B=\sum_{c=1}^{N-1}\beta_{c}e_{cN}$ and $A=e_{aN},$ and $A=e_{aa}$. We compute

@@

	\item $A=e_{aN}$ with $a\neq 1$

@@

\item $A=e_{aa}$ with $a\neq 1$
		\exp{\left(-\sum_{c=1}^{N-1}\beta_{c}e_{cN}^{x}\right)}e_{aa}^{x}\exp{\left(\sum_{c=1}^{N}\beta_{c}e_{cN}^{x}\right)}=e_{aa}^{x}+\beta_{a}e_{aN}^{x}

@@

Therefore, we have for the left boundary

@@

		e_{NN}^{1}-\mathbbm{1}^{1}+\sum_{a=1}^{N-1}(\alpha_{a}-\beta_{a})e_{aN}^{1}
Similarly, for the right boundary

@@

			e_{NN}^{L}-\mathbbm{1}^{L}
Obviously

@@

	H^{''}|\Psi^{''}\rangle=0
It is convenient to introduce also the whole transformation

@@

and, obviously

>>>>

In this section we aim to use duality and the matrix product ansatz (MPA) to determine the moments and correlations of the non-equilibrium steady state for the integrable version of the multi-species stirring process, defined in Subsection \ref{integrableChain-duality}. As it will turn out, the commutation relations of the matrix product ansatz allowing to determine a closed formula for all correlations. As a by-product, absorption probabilities can be explicitly found.%, giving a further probabilistic interpretation for the non-equilibrium steady state. 
\subsection{The integrable chain and the non-equilibrium steady state}

@@

 the master equation
    \frac{d|{\Psi}(t)\rangle}{dt}=H|{\Psi}(t)\rangle.

@@

	H|\Psi\rangle =0.

@@

	Z_{L}=\langle V^{*}|(X_{1}+\ldots +X_{N})^{L}|V\rangle.
Here the operators $(X_{a})_{a\in \{1,\ldots,N\}}$ fulfil the commutators

@@

	\langle V^{*}|\left(\alpha_{a}(X_{1}+\ldots+X_{N})-X_{a}\right)=\lambda_{a}\langle V^{*}|\qquad\forall a=1,\ldots,N,

@@

	\left(\beta_{a}(X_{1}+\ldots+X_{N})-X_{a}\right)|V\rangle=-\lambda_{a}|V\rangle\qquad\forall a=1,\ldots,N.

@@

\textbf{Remark}: Only $N-1$ of the $N$ equations \eqref{leftBoundary}  are independent. This can be seen by summing them over the species $a\in \{1,\ldots,N\}$, i.e.
		\sum_{a=1}^{N}	\langle V^{*}|\left(\alpha_{a}(X_{1}+\ldots+X_{N})-X_{a}\right)=0,%\sum_{a=1}^{N}\lambda_{a}\langle V^{*}|=0

@@

The MPA gives an abstract form of the steady state in terms of an algebra of operators. This path has been followed in \cite{vanicat2017exact} to find one and two-point correlations. 
Here we follow the same approach. We will show that the existence of absorbing dual process established in Section \ref{sectionDuality} simplifies substantially the  use of commutators in the

@@

where $H^{'}= S_{1} H S_{1}^{-1} $ has both boundary terms in a triangular form, and $H^{''} = S_{2}H' S_{2}^{-1}$ has the left boundary in a triangular form and the right boundary is diagonal. Using these transformations the commutators \eqref{bulk} and the action on the boundary vectors \eqref{leftBoundary}, \eqref{rightBoundary} become simpler. As a consequence, the steady state $|\Psi^{''}\rangle$ of $H^{''}$ can be written explicitly. Finally, reversing the transformations we can retrieve $|\Psi\rangle$.

The first transformation $S_{1}$ is closely related to the duality matrix and we will see that $H^{'}$ is $\widetilde{H}^{T}$, up to the extra-site term described by the bosonic creation operator in \eqref{siteDualHamiltonian}. Therefore, $H^{'}$ is not stochastic. The second transformation $S_{2}$ diagonalizes the right boundary. Finally, the composition  $S_{2} S_{1}$ allows to simplify the commutators of the MPA algebra. \\

The computation are organized as follows: we first define the similarity transformations $S_{1},S_{2}$ and the Hamiltonians $H^{'},H^{''}$. Second, we compute right eigenvector with vanishing eigenvalue of $H^{''}$ exploiting the simpler commutators. Finally, we invert the similarities and retrieve the right eigenvector of $H^{'}$ with vanishing eigenvalue and  the arbitrary correlations in the non-equilibrium steady state. The explicit expression of the steady state $|\Psi\rangle$ of $H$ is reported in Appendix \ref{appB}.

@@

	\mathcal{S}_{1}:=\exp{\left(\sum_{a=1}^{N-1}e_{Na}\right)}\label{S1Gen},\\
	\mathcal{S}_{2}:=\exp{\left(-\sum_{a=1}^{N-1}\beta_{a}e_{aN}\right)}.\label{S2Gen}

@@

 \mathcal{S}_{1}^{x}:=\exp{\left(\sum_{a=1}^{N-1}e_{Na}^{x}\right)},%=\mathbbm{1}+\sum_{a=2}^{N}e_{1a}

@@

 The transformation $S_{1}$ is invertible. We introduce 

@@

   H^{'}:=H^{'}_{left}+H_{bulk}+H^{'}_{right}

@@

	H^{'}_{left}=\sum_{a=1}^{N-1}\left(\alpha_{a}e_{aN}^{1}-e_{aa}^{1}\right)\qquad H^{'}_{right}=\sum_{a=1}^{N-1}\left(\beta_{a}e_{aN}^{L}-e_{aa}^{L}\right)

@@

		\sum_{a=1}^{N-1}\left(\alpha_{a}e_{aN}^{1}-e_{aa}^{1}\right).

@@

H_{right}^{'}=\exp{\left(\sum_{c=1}^{N-1}e_{1N}^{L}\right)}\sum_{a,b=1}^{N}\alpha_{a}\left(e_{ab}^{L}-e_{bb}^{L}\right)\exp{\left(-\sum_{c=1}^{N-1}e_{Nc}^{L}\right)}=\sum_{a=1}^{N-1}\left(\beta_{a}e_{aN}^{L}-e_{aa}^{L}\right)
\textbf{Remark:} The boundary Hamiltonians $H_{left}^{'}$ and $H_{right}^{'}$ relates to the transposed of  $\widetilde{H}_{left}$ and $\widetilde{H}_{right}$, i.e. to the transposed of the boundary part of the dual Hamiltonian  defined in \eqref{siteDualHamiltonian}. They are obtained replacing the  extra-site bosonic creation operators $(a^{\dagger})_{a}^{0}$  (resp. $(a^{\dagger})_{a}^{L+1}$) with the corresponding reservoir parameters $\alpha_a$ (resp. $\beta_a$).

@@

We have the following correspondence between eigenvectors with zero eigenvalues

@@

Therefore, by taking the product over the chain we obtain 

@@

	S_{2}=\prod_{x=1}^{L}S_{2}^{x}.

@@

	H^{''}=S_{2}H^{'}S_{2}^{-1}.

@@

\begin{align}\label{tri-diag-boundaryHamiltonians}
	H_{left}^{''}=e_{NN}^{1}-\mathbbm{1}+\sum_{a=1}^{N-1}(\alpha_{a}-\beta_{a})e_{aN}^{1}\qquad H_{right}^{''}=e_{NN}^{L}-\mathbbm{1}.
The left one is lower triangular and depends on the differences between the boundary parameters, while the right one is diagonal. Equations \eqref{tri-diag-boundaryHamiltonians} are proved using the Hadamard formula \eqref{HadamardFormula} with $B=\sum_{c=1}^{N-1}\beta_{c}e_{cN}$ and $A_{N}=e_{aN},$ and $A_{a}=e_{aa}$.

@@

	\item $A_{N}=e_{aN}$ with $a\neq 1$

@@

\item $A_{a}=e_{aa}$ with $a\neq 1$
		\exp{\left(-\sum_{c=1}^{N-1}\beta_{c}e_{cN}\right)}e_{aa}\exp{\left(\sum_{c=1}^{N}\beta_{c}e_{cN}\right)}=e_{aa}+\beta_{a}e_{aN}

@@

Therefore, using \eqref{HT-BA} we obtain for the left boundary 

@@

		e_{NN}^{1}-\mathbbm{1}+\sum_{a=1}^{N-1}(\alpha_{a}-\beta_{a})e_{aN}^{1}.
Similarly, using again \eqref{HT-BA} we have for the right boundary 

@@

			e_{NN}^{L}-\mathbbm{1}.
The relation between eigenvectors with vanishing eigenvalue is 

@@

	H^{''}|\Psi^{''}\rangle=0.
It is convenient to introduce also the transformation

@@

and, the relation between eigenvectors with vanishing eigenvalue is
